---
title: "Untitled"
author: "Julian Gradziel"
date: "2024-11-02"
output: pdf_document
---
```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(readr)
library(purrr)
library(tibble)
library(stringr)
library(lubridate)
library(forcats)
library(reshape2)
library(data.table)
library(caret)
library(randomForest)
library(xgboost)
library(shiny)
library(plotly)
library(leaflet)
library(RColorBrewer)
library(knitr)
library(rmarkdown)

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}
```{r}
library(data.table)

#--- 1) Set working directory
setwd("M:/OE0836/OE0709_shared/OVD/Christian - calibration")

#--- 2) Read first CSV and drop duplicates
df <- fread("ovd_EBOE.csv", sep=";", header=TRUE)
df <- unique(df)  # Removes duplicates

#--- 3) Rename columns to match your Python code
setnames(df,
         old  = c("T_BITA_COB_DT", "AMT_CUST_PRINC", "CCY_CUST_PRINC"),
         new  = c("cob_dt",        "AMT",            "cur"))

#--- 4) Convert date columns (adjust “tryFormats” to suit your data)
df[, cob_dt     := as.IDate(cob_dt,     tryFormats = c("%Y-%m-%d","%d/%m/%Y"))]
df[, DT_ACC_OPN := as.IDate(DT_ACC_OPN, tryFormats = c("%Y-%m-%d","%d/%m/%Y"))]

#--- 5) Convert AMT from “X,Y” to numeric
df[, AMT := as.numeric(gsub(",", ".", AMT))]

#--- 6) Compute the “first_dt_on” per GID_CONT (similar to groupby min)
dfint <- df[, .(first_dt_on = min(cob_dt, na.rm=TRUE)), by=GID_CONT]
df <- merge(df, dfint, by="GID_CONT", all.x=TRUE)

#--- 7) Now read the second dataset
df <- fread("4052_OVD_input.csv", sep=";")

#--- 8) Filter rows
df <- df[MODEL_ID == "4052_NP_OVD"]
df <- df[FLG_NPE == "N"]

#--- 9) Sort by cob_dt & cur
setorder(df, cob_dt, cur)

#--- 10) Sum AMT by group (like df.groupby(...).sum() in Python)
df <- df[ , .(AMT = sum(AMT, na.rm=TRUE)),
           by=.(cur, cob_dt, T_BITA_TNT_CD, GID_CONT,
                DT_ACC_OPN, FLG_NPE, MODEL_ID, first_dt_on)]
setorder(df, cur, cob_dt)

#--- 11) Pivot wider (like Pandas pivot_table)
index_cols <- c("cur", "T_BITA_TNT_CD", "GID_CONT",
                "DT_ACC_OPN", "FLG_NPE", "MODEL_ID", "first_dt_on")
df_wide <- dcast(df,
                 formula   = paste(paste(index_cols, collapse="+"), "~ cob_dt"),
                 value.var = "AMT")

#--- 12) Make a “previous” version shifted by 1 (df_prev_1 in Python)
df_wide_prev <- copy(df_wide)
value_cols   <- setdiff(names(df_wide), index_cols)

# Shift each date‐column of AMT by 1 so the first becomes NA
df_wide_prev[, (value_cols) := shift(.SD, n=1, fill=NA, type="lag"),
             .SDcols = value_cols]

#--- 13) Melt (long‐form) so we can do the same "stack" style as Python
df_prev_1 <- melt(df_wide_prev,
                  id.vars        = index_cols,
                  variable.name  = "cob_dt",
                  value.name     = "AMT_prev")
df_melted  <- melt(df_wide,
                   id.vars        = index_cols,
                   variable.name  = "cob_dt",
                   value.name     = "AMT")

# Convert cob_dt back to Date/IDate (it becomes character on melt)
df_prev_1[, cob_dt := as.IDate(cob_dt)]
df_melted[,  cob_dt := as.IDate(cob_dt)]

# Combine them by matching keys (cob_dt etc.)
df <- merge(df_melted, df_prev_1,
            by = c(index_cols, "cob_dt"),
            all.x = TRUE)

#--- 14) Replicate the Python difference columns
df[, Change_EoM    := fifelse(is.na(AMT), 0, AMT) -
                        fifelse(is.na(AMT_prev), 0, AMT_prev)]
df[, AMT_INCREASED := fifelse(Change_EoM > 0, Change_EoM, 0)]
df[, AMT_DECREASED := fifelse(Change_EoM < 0, Change_EoM, 0)]
df[, AMT_NB        := fifelse(cob_dt == first_dt_on, Change_EoM, 0)]
df[, AMT_closed    := fifelse(is.na(AMT) & !is.na(AMT_prev) & AMT_prev > 0,
                              AMT_prev, 0)]

#--- 15) Aggregate sums by (cur, cob_dt, MODEL_ID), exactly like Python
df_agg <- df[,
  .( AMT           = sum(AMT,           na.rm=TRUE),
     AMT_prev      = sum(AMT_prev,      na.rm=TRUE),
     Change_EoM    = sum(Change_EoM,    na.rm=TRUE),
     AMT_INCREASED = sum(AMT_INCREASED, na.rm=TRUE),
     AMT_DECREASED = sum(AMT_DECREASED, na.rm=TRUE),
     AMT_NB        = sum(AMT_NB,        na.rm=TRUE),
     AMT_closed    = sum(AMT_closed,    na.rm=TRUE)
  ),
  by = .(cur, cob_dt, MODEL_ID)
]

#--- 16) Sort (descending) by cob_dt if you like, as in Python (ascending=False)
setorder(df_agg, cur, -cob_dt)

#--- 17) Add check column
df_agg[, CheckChangeSum := AMT - AMT_prev]

#--- 18) Overwrite the first row's columns with 0 (mimicking df_agg.loc[0, ...] = 0)
df_agg[1, c("Change_EoM","AMT_INCREASED",
            "AMT_DECREASED","CheckChangeSum") := 0]

#--- 19) Finally, write out to CSV
fwrite(df_agg, "4052_NP_OVD_jupyter_new.csv", sep=";")


```
```

```{r}
library(data.table)

# 1) Set working directory
setwd("M:/OE0836/OE0709_shared/OVD/Christian - calibration")

#------------------------------------------------------------------------------
# 2) Read CSVs and remove duplicates
df       <- fread("ovd_SLSP.csv", sep=";", header=TRUE)
df       <- unique(df)
df_EURRON <- fread("EUR_RON.csv", sep=";", header=TRUE)
df_EURRON <- unique(df_EURRON)

#------------------------------------------------------------------------------
# 3) Filter to desired MODEL_ID & FLG_NPE (example: "2160_RON_P_OVD", "N")
df <- df[MODEL_ID == "2160_RON_P_OVD" & FLG_NPE == "N"]

#------------------------------------------------------------------------------
# 4) Rename columns to match your usage
#    Suppose your Python rename was something like:
#      t_bita_cob_dt   --> cob_dt
#      amt_cust_princ  --> AMT
#      ccy_cust_princ  --> cur
#      cluster         --> MODEL_ID
setnames(df,
         old  = c("t_bita_cob_dt","amt_cust_princ","ccy_cust_princ","cluster"),
         new  = c("cob_dt","AMT","cur","MODEL_ID"))

#------------------------------------------------------------------------------
# 5) Convert COB date column; keep only rows where cur == "EUR"
df[, cob_dt := as.IDate(cob_dt, tryFormats=c("%Y-%m-%d","%d/%m/%Y"))]
df <- df[cur == "EUR"]

#------------------------------------------------------------------------------
# 6) Sort by cob_dt, cur
setorder(df, cob_dt, cur)

#------------------------------------------------------------------------------
# 7) (Optional) Write out “raw” data if you like
fwrite(df, "2160_RON_P_OVD_rawdata.csv", sep=";")

#------------------------------------------------------------------------------
# 8) Create a “first_dt_on” by grouping on gid_cont
dfint <- df[, .(first_dt_on = min(cob_dt, na.rm=TRUE)), by=gid_cont]
df <- merge(df, dfint, by="gid_cont", all.x=TRUE)

#------------------------------------------------------------------------------
# 9) Create a “df_prev” with MonthEnd shifted date, plus a renamed MODEL_ID
df_prev <- copy(df)
# data.table does not have a direct “MonthEnd” offset built in.  One hack:
# shift cob_dt to the last day of the same month:
df_prev[, cob_dt := as.IDate(seq(as.Date(cob_dt), length.out=1, by="month")) - 1L]
# ^ adjust if you want the literal month-end date.  Another approach is to use lubridate.

# Rename model_id to track the “previous” version
setnames(df_prev, old="MODEL_ID", new="MODEL_ID_prev")

cluster <- "2140_CC"   # example from your code

#------------------------------------------------------------------------------
# 10) Aggregate the amounts in df (like groupby(...).sum()) to prepare for pivot
df_grouped <- df[, .(AMT = sum(AMT, na.rm=TRUE)),
                   by=.(cur, cob_dt, t_bita_tnt_cd, gid_cont, FLG_NPE, MODEL_ID, first_dt_on)]
setorder(df_grouped, cur, cob_dt)

#------------------------------------------------------------------------------
# 11) Pivot to wide format (like pandas pivot_table)
#     This will spread each cob_dt into a column of AMT
df_wide <- dcast(
  df_grouped,
  formula   = cur + t_bita_tnt_cd + gid_cont + FLG_NPE + MODEL_ID + first_dt_on ~ cob_dt,
  value.var = "AMT",
  fill      = 0
)

#------------------------------------------------------------------------------
# 12) Create a “previous wide” dataset by shifting columns one to the right 
#     (similar to your Python code that loops and assigns df_prev_1[i] = df[i-1])
df_wide_prev <- copy(df_wide)
wide_cols <- setdiff(names(df_wide), c("cur","t_bita_tnt_cd","gid_cont","FLG_NPE","MODEL_ID","first_dt_on"))
# Shift each date-column to “simulate” prior-month values
df_wide_prev[, (wide_cols) := shift(.SD, n=1, fill=NA, type="lag"), .SDcols=wide_cols]

#------------------------------------------------------------------------------
# 13) Convert back to “long” form so we can merge
df_long      <- melt(df_wide,
                     id.vars       = c("cur","t_bita_tnt_cd","gid_cont","FLG_NPE","MODEL_ID","first_dt_on"),
                     variable.name = "cob_dt",
                     value.name    = "AMT")
df_prev_long <- melt(df_wide_prev,
                     id.vars       = c("cur","t_bita_tnt_cd","gid_cont","FLG_NPE","MODEL_ID","first_dt_on"),
                     variable.name = "cob_dt",
                     value.name    = "AMT_prev")

# Convert “cob_dt” from character to Date or IDate
df_long[,      cob_dt := as.IDate(cob_dt)]
df_prev_long[, cob_dt := as.IDate(cob_dt)]

# Merge them together
df_final <- merge(df_long, df_prev_long,
                  by = c("cur","t_bita_tnt_cd","gid_cont","FLG_NPE","MODEL_ID","first_dt_on","cob_dt"),
                  all.x = TRUE)

#------------------------------------------------------------------------------
# 14) Fill NA with 0, compute difference columns, etc.
df_final[is.na(AMT),      AMT := 0]
df_final[is.na(AMT_prev), AMT_prev := 0]

df_final[, Change_EoM    := AMT - AMT_prev]
df_final[, AMT_INCREASED := fifelse(Change_EoM > 0, Change_EoM, 0)]
df_final[, AMT_DECREASED := fifelse(Change_EoM < 0, Change_EoM, 0)]
df_final[, AMT_NB        := fifelse(cob_dt == first_dt_on, Change_EoM, 0)]
df_final[, AMT_closed    := fifelse(AMT == 0 & AMT_prev > 0, AMT_prev, 0)]

#------------------------------------------------------------------------------
# 15) Suppose you want to group and sum these final columns by (cur, cob_dt, MODEL_ID),
#     mimicking your “df.groupby([...])[...].sum()”.
df_agg <- df_final[,
  .( AMT           = sum(AMT,           na.rm=TRUE),
     AMT_prev      = sum(AMT_prev,      na.rm=TRUE),
     Change_EoM    = sum(Change_EoM,    na.rm=TRUE),
     AMT_INCREASED = sum(AMT_INCREASED, na.rm=TRUE),
     AMT_DECREASED = sum(AMT_DECREASED, na.rm=TRUE),
     AMT_NB        = sum(AMT_NB,        na.rm=TRUE),
     AMT_closed    = sum(AMT_closed,    na.rm=TRUE)
   ),
   by=.(cur, cob_dt, MODEL_ID)
]
# Sort descending by cob_dt if desired
setorder(df_agg, cur, -cob_dt)

#------------------------------------------------------------------------------
# 16) Optionally, zero out the first row’s difference columns (like .loc[0, ...] in Python)
df_agg[1, c("Change_EoM","AMT_INCREASED","AMT_DECREASED") := 0]
df_agg[1, CheckChangeSum := 0]

#------------------------------------------------------------------------------
# 17) Write aggregated results
fwrite(df_agg, "2140_OVD_aggregated_no_change_1.csv", sep=";")

#------------------------------------------------------------------------------
# 18) Other filters or final checks: for example,
#     slicing df_final for certain cob_dt dates
dftest <- df_final[
  cob_dt %in% as.IDate(c("2024-07-31","2024-08-31")) &
  !is.na(AMT)
]
print(dftest)

# or group again for a final summary
df_agg_changed <- df_final[, 
  .(AMT            = sum(AMT, na.rm=TRUE),
    AMT_prev       = sum(AMT_prev, na.rm=TRUE),
    Change_EoM     = sum(Change_EoM, na.rm=TRUE),
    AMT_INCREASED  = sum(AMT_INCREASED, na.rm=TRUE),
    AMT_DECREASED  = sum(AMT_DECREASED, na.rm=TRUE),
    AMT_NB         = sum(AMT_NB, na.rm=TRUE),
    AMT_closed     = sum(AMT_closed, na.rm=TRUE)),
  by=.(cur, cob_dt, MODEL_ID)
]
setorder(df_agg_changed, MODEL_ID, cob_dt)
print(df_agg_changed)

```

```{r}
library(data.table)
library(lubridate)  # Optional for month-end manipulations

#------------------------------------------------------------------------------
# 1) Read & de-duplicate 
#    (In Python: df = pd.read_csv('ovd_BCR.csv', sep=';').drop_duplicates())
#------------------------------------------------------------------------------
df       <- fread("ovd_BCR.csv", sep=";", header=TRUE)
df       <- unique(df)
df_EURRON <- fread("EUR_RON.csv", sep=";", header=TRUE)
df_EURRON <- unique(df_EURRON)

#------------------------------------------------------------------------------
# 2) Filter to desired MODEL_ID & FLG_NPE (like df[df["MODEL_ID"]=="2160_RON_P_OVD"]
#    & df[df["FLG_NPE"]=="N"])
#------------------------------------------------------------------------------
df <- df[MODEL_ID == "2160_RON_P_OVD" & FLG_NPE == "N"]

#------------------------------------------------------------------------------
# 3) Rename columns 
#    (In Python: df.rename(columns={'T_BITA_COB_DT':'cob_dt','AMT_CUST_PRINC':'AMT','CCY_CUST_PRINC':'cur'}, ...))
#------------------------------------------------------------------------------
setnames(df,
         old=c("T_BITA_COB_DT","AMT_CUST_PRINC","CCY_CUST_PRINC"),
         new=c("cob_dt","AMT","cur"))

# Rename the FX column in df_EURRON if needed
# (In Python: df_EURRON.rename(columns={'RT_FX_FIX':'FX_EURRON'}, ...))
setnames(df_EURRON,
         old="RT_FX_FIX",
         new="FX_EURRON")

#------------------------------------------------------------------------------
# 4) Parse date columns
#    (In Python: df["cob_dt"] = pd.to_datetime(df["cob_dt"], infer_datetime_format=True))
#------------------------------------------------------------------------------
df[, cob_dt         := as.IDate(cob_dt,         tryFormats=c("%Y-%m-%d","%d/%m/%Y"))]
df[, DT_ACC_OPN     := as.IDate(DT_ACC_OPN,     tryFormats=c("%Y-%m-%d","%d/%m/%Y"), optional=TRUE)]
df_EURRON[, T_BITA_COB_DT := as.IDate(T_BITA_COB_DT, tryFormats=c("%Y-%m-%d","%d/%m/%Y"))]

#------------------------------------------------------------------------------
# 5) Replace commas in numeric columns, convert to numeric
#    (In Python: df["AMT"] = df["AMT"].str.replace(',','.').astype(float))
#------------------------------------------------------------------------------
df[, AMT := as.numeric(gsub(",", ".", AMT))]
df_EURRON[, FX_EURRON := as.numeric(gsub(",", ".", FX_EURRON))]

#------------------------------------------------------------------------------
# 6) Merge the FX table on matching date 
#    (In Python: df = df.merge(df_EURRON[["T_BITA_COB_DT","FX_EURRON"]],
#                              left_on="cob_dt", right_on="T_BITA_COB_DT"))
#------------------------------------------------------------------------------
# Sort if needed (like df_EURRON.sort_values in Python):
setorder(df_EURRON, T_BITA_COB_DT)
df <- merge(df, 
            df_EURRON[, .(T_BITA_COB_DT, FX_EURRON)], 
            by.x="cob_dt", by.y="T_BITA_COB_DT",
            all.x=TRUE)

# Convert the amounts (In Python: df["AMT"] = df["AMT"] / df["FX_EURRON"])
df[, AMT := AMT / FX_EURRON]

#------------------------------------------------------------------------------
# 7) Derive “first_dt_on” by GID_CONT for any positive AMT
#    (In Python: groupby min date where AMT>0)
#------------------------------------------------------------------------------
dfint <- df[AMT > 0, .(first_dt_on = min(cob_dt, na.rm=TRUE)), by=GID_CONT]
df <- merge(df, dfint, by="GID_CONT", all.x=TRUE)

#------------------------------------------------------------------------------
# 8) Summarize the amounts by grouping, as your Python code does:
#    df.groupby(["cur","cob_dt","T_BITA_TNT_CD","GID_CONT","DT_ACC_OPN","FLG_NPE","MODEL_ID","first_dt_on"])["AMT"].sum()
#------------------------------------------------------------------------------
df_summarized <- df[,
  .(AMT = sum(AMT, na.rm=TRUE)),
  by=.(cur, cob_dt, T_BITA_TNT_CD, GID_CONT, DT_ACC_OPN, FLG_NPE, MODEL_ID, first_dt_on)
]
setorder(df_summarized, cur, cob_dt)

#------------------------------------------------------------------------------
# 9) Pivot to wide (like df.pivot(index=..., columns="cob_dt", values="AMT"))
#    in data.table we use dcast
#------------------------------------------------------------------------------
index_cols <- c("cur","T_BITA_TNT_CD","GID_CONT","DT_ACC_OPN","FLG_NPE","MODEL_ID","first_dt_on")
df_wide <- dcast(
  df_summarized,
  formula = paste(paste(index_cols, collapse="+"), "~ cob_dt"),
  value.var = "AMT",
  fill = 0
)

#------------------------------------------------------------------------------
# 10) Make a copy, shift each “date column” by 1 to create the “previous” columns
#    (In Python: for i in range(1, len(df.columns)): df_prev[i] = df[i-1])
#------------------------------------------------------------------------------
df_prev_wide <- copy(df_wide)

# Identify the pivoted date columns
wide_date_cols <- setdiff(names(df_wide), index_cols)

# Shift them by 1 (so col[i] = col[i-1], the first col becomes NA)
df_prev_wide[, (wide_date_cols) := shift(.SD, n=1, fill=NA, type="lag"),
             .SDcols=wide_date_cols]

#------------------------------------------------------------------------------
# 11) Convert both wide tables back to long form so we can join “AMT” and “AMT_prev”
#    (In Python: df = df.stack(...).reset_index(), etc.)
#------------------------------------------------------------------------------
df_long <- melt(df_wide,
                id.vars=index_cols,
                variable.name="cob_dt",
                value.name="AMT")

df_prev_long <- melt(df_prev_wide,
                     id.vars=index_cols,
                     variable.name="cob_dt",
                     value.name="AMT_prev")

# Convert cob_dt from character to Date/IDate
df_long[, cob_dt := as.IDate(cob_dt)]
df_prev_long[, cob_dt := as.IDate(cob_dt)]

# Merge them
df_final <- merge(df_long, df_prev_long,
                  by=c(index_cols, "cob_dt"),
                  all.x=TRUE)

#------------------------------------------------------------------------------
# 12) Fill NA with 0, compute difference columns
#    (In Python: df["AMT"] = df["AMT"].fillna(0), etc.)
#------------------------------------------------------------------------------
df_final[is.na(AMT),      AMT := 0]
df_final[is.na(AMT_prev), AMT_prev := 0]

df_final[, Change_EoM    := AMT - AMT_prev]
df_final[, AMT_INCREASED := fifelse(Change_EoM > 0, Change_EoM, 0)]
df_final[, AMT_DECREASED := fifelse(Change_EoM < 0, Change_EoM, 0)]

# Example logic for “AMT_NB”:
# (In Python you had a condition using DT_ACC_OPN + offsets.MonthEnd() >= cob_dt, etc.)
# Let’s assume you want: if cob_dt == first_dt_on => “new business”
df_final[, AMT_NB := fifelse(cob_dt == first_dt_on, Change_EoM, 0)]

# For “AMT_closed” (like df.where(AMT==0 & AMT_prev>0, AMT_prev, 0))
df_final[, AMT_closed := fifelse(AMT == 0 & AMT_prev > 0, AMT_prev, 0)]

#------------------------------------------------------------------------------
# 13) Final aggregation (like Python’s df.groupby(...).sum().reset_index())
#------------------------------------------------------------------------------
df_agg <- df_final[,
  .( AMT           = sum(AMT,           na.rm=TRUE),
     AMT_prev      = sum(AMT_prev,      na.rm=TRUE),
     Change_EoM    = sum(Change_EoM,    na.rm=TRUE),
     AMT_INCREASED = sum(AMT_INCREASED, na.rm=TRUE),
     AMT_DECREASED = sum(AMT_DECREASED, na.rm=TRUE),
     AMT_NB        = sum(AMT_NB,        na.rm=TRUE),
     AMT_closed    = sum(AMT_closed,    na.rm=TRUE)
   ),
   by=.(cur, cob_dt, MODEL_ID)
]
setorder(df_agg, cur, -cob_dt)

#------------------------------------------------------------------------------
# 14) Optionally zero out the first row’s difference columns (like df_agg.loc[0, ...] = 0)
#------------------------------------------------------------------------------
if (nrow(df_agg) > 0) {
  df_agg[1, c("Change_EoM","AMT_INCREASED","AMT_DECREASED") := 0]
  df_agg[1, CheckChangeSum := 0]
}

#------------------------------------------------------------------------------
# 15) Export final results
#    (In Python: df_agg.to_csv('2160_RON_P_OVD_jupyter_RON_new.csv', sep=';'))
#------------------------------------------------------------------------------
fwrite(df_agg[cur=="RON"], "2160_RON_P_OVD_jupyter_RON_new.csv", sep=";")

#------------------------------------------------------------------------------
# 16) Example of date offset logic for “DT_ACC_OPN” + MonthEnd
#    In Python: df["DT_ACC_OPN"] + pd.offsets.MonthEnd()
#    In R, using lubridate: floor_date(x, "month") + months(1) - 1
#    or there is also a ceiling_date/floor_date approach. E.g.:
#------------------------------------------------------------------------------
df_final[, DT_ACC_OPN_plusMonthEnd := (floor_date(DT_ACC_OPN, "month") + months(1) - days(1))]

```



